Using Planning to Generate Natural-Language Sentences:\\ Experiences and Challenges

### 2009-12-18 trunk/Papers/crisp-ff-10/abstract.tex
skipped last sentence
shorter, new title

\begin{abstract}
We present a planning domain that encodes the problem of generating
natural language sentences. This domain has a number of features that
make it significantly different from existing IPC benchmark
domains. In particular, we show how several techniques used in the FF
planner, which has been highly successful on IPC problems, are led
astray. We discuss possible fixes to FF's shortcomings, some of which
we have already implemented, but some of which pose interesting
challenges to the AI Planning community. The performance of our
modified FF is up to three orders of magnitude better than that of the
original FF. While this makes FF much more practical for this
application, we show that it is still very limited with respect to
problem parameters such as the complexity of the underlying
grammar. We outline how other planning domains might share the
features that make the generation domain difficult. {\bf do we? how?}
\end{abstract}

>>>>

\begin{abstract}
We present a planning domain that encodes the problem of generating
natural language sentences. This domain has a number of features that
provoke fairly unusual behavior in existing planners. In particular,
we show how several techniques used in the FF planner, which has been
highly successful on IPC problems, are led astray. We discuss possible
fixes to FF's shortcomings, some of which we have already implemented,
but some of which pose interesting challenges to the AI Planning
community. The performance of our modified FF is up to several orders
of magnitude better than that of the original FF. While this makes FF
much more practical for this application, we show that it is still
quite limited with respect to the complexity of the underlying
grammar.
%We outline how other planning domains might share the
%features that make the generation domain difficult. {\bf do we? how?}
\end{abstract}

### 2009-12-18 trunk/Papers/crisp-ff-10/introduction.tex
shorter, new title

Natural language generation (NLG) -- the problem of computing a
sentence or text that communicates a given piece of information -- has
a long-standing connection to automated planning
\cite{perrault80,appelt:planning}. In the past few years, the idea has
been picked up again in the NLG community, both to use planning as an
approach for modeling language problems
\cite{Steedman-Petrick:07,benotti08b} and to tackle the search
problems in existing NLG problems by using techniques from planning
\cite{KolSto07}.

Existing planners are relatively good at solving modest instances of
the planning problems that arise in this context. However, Koller and
Petrick \shortcite{koller10:_exper_with_plann_for_natur_languag_gener}
recently showed that as the planning problem instances from Koller and
Stone \shortcite{KolSto07} scale up, off-the-shelf planners become too
slow to be practically useful. They noted that there are major sources
of inefficiency both in the search itself and in the commonly used
preprocesses, which are put under pressure by several features of the
domain (e.g.\ comparatively large predicate and operator arities) that
diverge from most IPC benchmarks.
%the sheer number of
%ground instances of facts and actions that come up in this domain.

In this paper, we revisit this situation, analyze the sources of
inefficiency in more detail in the context of the FF planner
\cite{HoffmannNebel01}, and show how FF can be modified to perform
dramatically better. We achieve this by tweaking the implementation of
the FF preprocessor, and by acting on some simple but shocking
observations regarding the behavior of some of FF's search techniques,
in particular its goal ordering technique and the relaxed plan
heuristic. The changes we make are all simple, but make FF perform
well enough to be useful in practical NLG applications.

@@

Our performance gains with simple FF configuration changes certainly
point out that automatic planner configuration may be quite
useful. Apart from that, our domain remains an interesting challenge
for the development of better heuristics. There still are realistic
instances of the NLG planning domain on which our modified FF becomes
too slow for practical use. Overcoming this weakness would entail
overcoming the observed weaknesses of FF's search techniques --
variants of which are used in many planners today.

>>>>

Natural language generation (NLG) -- the problem of computing a
sentence or text that communicates a given piece of information -- has
a long-standing connection to automated planning
\cite{perrault80,appelt:planning}. In the past few years, the idea has
been picked up again in the NLG community, both to use planning as an
approach for modeling language problems
\cite{Steedman-Petrick:07,benotti08b} and to tackle the search
problems in NLG by using techniques from planning \cite{KolSto07}.

Existing planners are relatively good at solving modest instances of
the planning problems that arise in this context. However, Koller and
Petrick \shortcite{koller10:_exper_with_plann_for_natur_languag_gener}
recently showed that as the planning problem instances from Koller and
Stone \shortcite{KolSto07} scale up, off-the-shelf planners become too
slow to be practically useful. They noted that there are major sources
of inefficiency both in the search itself and in the commonly used
preprocesses, due to several features of the domain
(e.g.\ comparatively large predicate and operator arities) that
diverge from most IPC benchmarks.
%the sheer number of
%ground instances of facts and actions that come up in this domain.

In this paper, we revisit this situation, analyze the sources of
inefficiency in more detail in the context of the FF planner
\cite{HoffmannNebel01}, and show how to ``wake up the rabbit'': we
modify FF to obtain dramatically better performance. We achieve this
by tweaking the implementation of FF's preprocessor, and by acting on
some simple but shocking observations regarding the behavior of some
of FF's search techniques, in particular its goal ordering technique
and the relaxed plan heuristic. The changes we make are all simple,
but make FF perform well enough to be useful in practical NLG
applications.

@@

Our performance gains with simple FF configuration changes certainly
point out that automatic planner configuration may be quite
useful. Apart from that, our domain remains an interesting challenge
for the development of better heuristics. There still are realistic
instances of the NLG planning domain on which our modified FF is not
practical. Overcoming this weakness would entail overcoming the
observed weaknesses of FF's search techniques -- variants of which are
used in many planners today.
